---
title: "Trying to add data and rmeasure to POMP model"
output: html_notebook
---
  This is a baby attempt to include data and sampling process with likelihood function.

  I am using the most basic closed stochastic SIR model, with initial states fixed.
  I will want to fit gamma and beta.
  I will be incuding data generated to more or less match gamma = 1/13, beta = 1 (unit is days here).
  The data measures the number of positive individuals out of a sample of 100.
  I want to see if it finds original parameters if it gets good data, to check whether my approach works at all.

  This s the original POMP model:
```{r pomp_model}
library(pomp)
# stochastic discrete SIR model in POMP
# no pop dyn.

# stochastic stepping process
stochStep <- Csnippet("
  p_SI = 1 - exp(-Beta*I/N);
  p_IR = 1 - exp(-gamma);

  n_SI = rbinom(S,p_SI);
  n_IR = rbinom(I,p_IR);

  S = S - n_SI;
  I = I + n_SI - n_IR;
  R = R + n_IR;

")

# deterministic skeleton
skel <- Csnippet("
  DS = S - Beta*S*I/N;
  DI = I + Beta*S*I/N-gamma*I;
  DR = R + gamma*I;
")

# initial values of states
init1 <- Csnippet("
  S = N-5;
  I = 5;
  R = 0;
  ")

params1 <- c(Beta=1, gamma = 1/13, N= 1000)


closed_sir <- pomp(data=data.frame(time=1:51,data=NA),
                   times="time",
                   t0=0,
                   rprocess=discrete_time(step.fun=stochStep,delta.t=1),
                   skeleton = map(skel),
                   rinit=init1,
                   statenames=c("S","I","R", "p_SI", "p_IR", "n_SI", "n_IR"),
                   paramnames=c("Beta","gamma","N"))




sim <- simulate(closed_sir,params=params1,format = "data.frame", nsim = 100)

traj <- trajectory(closed_sir,params=params1,format="data.frame")

```
Let's plot the model to see what we are working with:
``` {r plot}
ggplot(NULL)+
geom_line(data=sim, alpha = 0.05,aes(x=time, y = I, group= factor(.id), colour = "Stochastic infected"))+
geom_line(data=sim, alpha = 0.05,aes(x=time, y = R, group= factor(.id), colour = "Stochastic recovered"))+
geom_line(data=sim, alpha = 0.05,aes(x=time, y = S, group= factor(.id), colour = "Stochastic susceptible"))+
scale_color_discrete(name="Legend")+
geom_line(data=traj, aes(x=time, y=I))+
geom_line(data=traj, aes(x=time, y=R))+
geom_line(data=traj, aes(x=time, y=S))+
labs(title="Stochastic SIR model")
```
  
  Now we try to generate observations. We calculate the prevalence of the deterministic model, add a random normally distributed number with mean 0, sd 5, and round it to a non-negative integer to include some stochasticity.
  
```{r making the data}

traj %>% 
  filter(time %in% seq(1,51,by=10)) -> traj_filter

floor(traj_filter$I*100/(traj_filter$S+traj_filter$I+traj_filter$R) ) -> traj_filter$prev
pmax(0, floor(traj_filter$prev + rnorm(0,5, n=6) )) -> traj_filter$pos

traj_filter %>% select(time, pos) -> test_data

test_data
```
So now we have our data saved in test_data, the number of infecteds out of 100 samples, approximately matching the model's infection prevalence.  

  We can now write rmeasure and dmeasure, the first one is to draw, the second is to the likelihood of p(observation | current state of system and parameters).

``` {r measuring process}
rmeas <- Csnippet("pos = rbinom(100, I/N);")
dmeas <- Csnippet("lik = dbinom(pos, 100, I/N, give_log);")
```

Now we re-specify the POMP model to include our data and the measurement process:
``` {r re-specify model}

trial_model <- pomp(data=test_data,
                   times="time",
                   t0=0,
                   rprocess=discrete_time(step.fun=stochStep,delta.t=1),
                   skeleton = map(skel),
                   rinit=init1,
                   statenames=c("S","I","R", "p_SI", "p_IR", "n_SI", "n_IR"),
                   paramnames=c("Beta","gamma","N"),
                   rmeasure = rmeas,
                   dmeasure = dmeas
                   )
```

We now simulate the number of positives and plot the actual number of positives we generated:
``` {r simulate positives}

sim <- simulate(trial_model, params=params1, format = "data.frame", nsim = 5, ,include.data=TRUE)

ggplot(data=sim,
       mapping=aes(x=time,y=pos))+
  geom_line()+
  facet_wrap(~.id,ncol=1,scales="free_y")
```

  We can see that in general the simulations followed the same shape. This confirms that rmeasure is working, we can simulate things we wanted! 
  We can estimate the likelihood using a particle filter. params_good is the orignal value we used to create the data, so it should be a good fit. params_bad is an arbitrary set of parameters, to see if it gives us a worse likelihood. We print out the likelihoods of the 10 replications, and the plots of the effective sample size and conditional log likelihoods.
```{r}
params_good <- c(Beta=1, gamma = 1/13, N= 1000)
pf <- replicate(10, pfilter(trial_model, params = params_good ,Np=5000))

ll <- sapply(pf,logLik)
logmeanexp(ll,se=TRUE)

plot(pf[[1]])


params_bad <-c(Beta=0.1, gamma = 1/20, N= 5000)
pf2 <- replicate(10, pfilter(trial_model, params = params_bad ,Np=5000))

ll <- sapply(pf2,logLik)
logmeanexp(ll,se=TRUE)

plot(pf2[[1]])

```
  This confirms that dmeasure is also working, our fit is pretty good! Now we try some fitting methods, like iterated particle filtering to improve the bad set of parameters:
  
  THIS IS NOT WORKING, SORT THIS OUT:
```{r}
mif2(trial_model, params = params_bad, Nmif=30, Np=1000, 
     cooling.fraction.50=0.8,cooling.type="geometric",
     rw.sd=rw.sd(Beta=0.02,gamma=0.02,N=0.02)
) -> mf
plot(mf)
```