---
title: "Seasonal births SIR POMP"
author: "Alexandra Souly"
date: "07/07/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r load packages, include=FALSE}
library(pomp)
library(ggplot2)
```

I used POMP to create a deteministic and a stochastic version of the basic SIR model. I am using the discrete difference equations version for both. I tried adding in simple birth-death dynamics too, with yearly births accordings to $b(t)=b_0\,(1+b_1\cos(2\,\pi\,t))$.

POMP stores all data in its own pomp class that can be used for complicated methods in the package, so I hope it will be easier to implement algorithms later within the package.

We start by specifying the deterministic model skeleton, I use the following equations:  
  ("D" signifies the update value for the map in the equations)   
```{r skel}
skel <- Csnippet("
  double b = b0*(1+b1*cos(2*M_PI*t/365));
  DS = S - Beta*S*I/N           + b*N - d*S ;
  DI = I + Beta*S*I/N - gamma*I       - d*I;
  DR = R              + gamma*I       - d*R;
")
```
  Then I specify the stochastic stepping process:  
  I used 1-exp(r) for the per capita stepping probabilities for rate r in the difference eqs and a binomial distribution for the numbers of individuals moving between compartments.

```{r SthcStep}

stochStep <- Csnippet("
  double b = b0*(1+b1*cos(2*M_PI*t/365));
  p_SI = 1 - exp(-Beta*I/N);
  p_IR = 1 - exp(-gamma);
  p_b  = 1 - exp(-b);
  p_d  = 1 - exp(-d);

  n_SI = rbinom(S,p_SI);
  n_IR = rbinom(I,p_IR);
  
  b_N  = rbinom(N, p_b);
  d_I  = rbinom(I, p_d);
  d_R  = rbinom(R, p_d);
  d_S  = rbinom(S, p_d);
  
  S = S - n_SI        + b_N  - d_S;
  I = I + n_SI - n_IR        - d_I;
  R = R        + n_IR        - d_R;

")
```
I set some arbitrary initial states and parameters (5 infected out of, 13 days infectious period)

```{r setting initial values and params}
# initial values of states
init1 <- Csnippet("
  S = N-5;
  I = 5;
  R = 0;
  ")

params1 <- c(Beta=1, b0=0.1, b1=0.4, gamma = 1/13, N= 763, d = 0.1)
```

Then we can create a POMP object:  
  We follow the process for 700 days, POMP will use our stochastic process described above.

```{r POMP object}

closed_sir <- pomp(data=data.frame(time=1:700,data=NA),
                   times="time",
                   t0=0,
                   rprocess=discrete_time(step.fun=stochStep,delta.t=1),
                   skeleton = map(skel),
                   rinit=init1,
                   statenames=c("S","I","R", "p_SI", "p_IR", "p_b", "p_d", "n_SI", "n_IR", "b_N", "d_I", "d_S", "d_R"),
                   paramnames=c("Beta","b0","b1","gamma","N", "d"))

```

We then simulate 100 instances of the stochastic model and also calculate the trajectory with the deterministic model:  

``` {r simulate}
sim <- simulate(closed_sir,params=params1,format = "data.frame", nsim = 100)

traj <- trajectory(closed_sir,params=params1,format="data.frame")
```
We first plot the population dynamics to see how the seasonal birth pulse looks like:

```{r population plot}

birth_plot <- ggplot(NULL)+ geom_line(data=traj,aes(x=time, y = I+S+R,))+ylim(0,NA)+
              labs(title="Seasonal birth pulse")+xlab("Days")+ylab("Size of population")+
              theme(plot.title = element_text(hjust = 0.5))

plot(birth_plot)


```
We then plot to see how the stochastic results differ from the deterministic trajectory (in black). Red is infected, green is recovered and blue is supsceptible:


```{r Creating plot, echo=FALSE}

cols <- c( "Stochastic infected"= "red", "Stochastic susceptible" = "blue", "Stochastic recovered" = "darkgreen")

created_plot <-ggplot(NULL)+
  geom_line(data=sim, alpha = 0.05,aes(x=time, y = I, group= factor(.id), colour = "Stochastic infected"))+
  geom_line(data=sim, alpha = 0.05,aes(x=time, y = R, group= factor(.id), colour = "Stochastic recovered"))+
  geom_line(data=sim, alpha = 0.05,aes(x=time, y = S, group= factor(.id), colour = "Stochastic susceptible"))+
  scale_colour_manual(values = cols)+
  geom_line(data=traj, aes(x=time, y=I))+
  geom_line(data=traj, aes(x=time, y=R))+
  geom_line(data=traj, aes(x=time, y=S))+
  labs(title="Stochastic SIR model")

plot(created_plot)
```
