---
title: "Adding in measuring"
output:
  html_document:
    df_print: paged
---

The function I am using to simulate tests:

```{r rmeasure}
library(pomp)
  rmeas <- Csnippet("
                          
                          double N = (Sn + Sj + Sm + Sf + En + Ej + Em + Ef +                                       In + Ij + Im + If 
                                      + Rn + Rj + Rm + Rf + Ma);
                          double I_total =  In + Ij + Im + If;
                          double p = zeta * I_total / N;
                          double p_pool = 1 - pow((1 - p), d);
                          pos = rbetabinom(samplesize, p_pool, disp
                          ")
```
  For now, I used the following parameters that should be the posteriors Aaron got in his model 5, SIR with no seasonal forcing:
```{r params}
params <- c(
  
  # base dynamics params
  beta_val = 0.06, # infection rate S -> R
  gamma_val = 3.135, # recovery rate I -> R
  omega_val = 0, # immune waning rate R -> S
  omega_m_val = 0.8, # maternal antibody waning rate
  kappa_val = 2350, # carrying capacity
  rho_val = 0, # I -> L 
  epsilon_val = 0, # L -> I 
  mu_val = 1.37, # juvenile maturation rate
  mj_val = 0.50, # juvenile death rate
  m_val = 0.23, # adult death rate
  delta_t = 365, # scaling time as days instead of years
  
  # seasonality params
  c = 18.65, # birth pulse scaling factor
  s = 130, # birth pulse synchronicity
  phi = 7.19, # birth pulse time shift
  
  # measuring process params
  zeta = 0.3, # test accuracy
  disp = 1000, # dispersion parameter
  d = 10 # number of bats contributing to the same pool
)
``` 

 And the following equilibrium initial values, which I have got after 50 years of running:
```{r ini}
   init_states <- Csnippet("
                        
                        Ma = 501;
                        Sn = 3;
                        Sj = 70;
                        Sm = 3;
                        Sf = 3;
                        
                        En = 0;
                        Ej = 0;
                        Em = 0;
                        Ef = 0;
                        
                        In = 3;
                        Ij = 60;
                        Im = 7;
                        If = 7;
                        
                        Rn = 4;
                        Rj = 255;
                        Rm = 479;
                        Rf = 479;
                        
                        ")                       
```

Now we can see the model with 100 simulations:

```{r sims}
library(ggpubr)
source("full_pomp_model.R")
ggarrange(plotlist =plots, ncol = 3)
```

This looks as expected, I think the model is working fine. Now lets look at our simulated test results:
```{r}
plprev
```